|

### 2020-02-07 02-Chp2-Systematic-Review.Rmd
Major updates

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:darwin-quote)'}
There is grandeur in this view of life, with its several powers, having been originally breathed into a few forms or into one; and that, whilst this planet has gone cycling on according to the fixed law of gravity, from so simple a beginning endless forms most beautiful and most wonderful have been, and are being, evolved.
```
(ref:darwin-quote) --- Charles Darwin [@Darwin1859]

@@

## Test

>>>>

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:darwin-quote)'}
Hold
```
(ref:darwin-quote) --- Charles Darwin [@chang2012a]

@@

```{r, echo = FALSE}
knitr::read_chunk("scripts/Chp2-Systematic-Review.R")
```

## Aims

The aim of this chapter is to systematically review all available literature on the association between blood levels of total cholesterol and it's constituent parts (HDL-c,LDL-c and triglycerides) on the subsequent risk of dementia.

## Methods

### Search strategy

We will systematically search electronic bibliographic databases to identify potentially relevant records. The search strategy used in each database will be developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE) terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy will include terms related to lipids, lipid modifying treatments, and dementia and its sub-types, and will be designed for MEDLINE before being adapted for use in the other bibliography databases listed. An outline of the general strategy is presented in the Table 3.2 below and the full draft search strategies for each database are attached to this protocol. To ensure that the study design filters are not excluding potentially relevant records, a random sample of 500 records identified by the main search but excluded by the filters (defined as Line 7 NOT Line 13 in Table 3.2) will be screened. If any potentially relevant studies are identified, their titles and abstracts will be searched for key terms that can be incorporated into the filters to improve search sensitivity.

The following databases will be searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection. We will also search clinical trial registries, for example ClinicalTrials.gov, to identify relevant randomized controlled trials.

The abstracts list of relevant conferences (e.g. the proceedings of the Alzheimer’s Association International Conference, published in the journal Alzheimer’s & Dementia) will be searched. Grey literature will also be searched via ProQuest, OpenGrey and Web of Science Conference Proceedings Citation Index, while theses will be accessed using the Open Access Theses and Dissertations portal. We will also search BioArxiv, a preprint repository, to identify potentially relevant studies. Finally, the reference lists of included studies will be searched by hand while studies citing included studies will be examined using Google Scholar (forward and reverse citation searching).

### Study selection

Records will be imported into Endnote and deduplicated using the method outlined in Bramer et al. (2016).[@bramer2016] Screening (both title/abstract and full-text) will be performed using a combination of Endnote and Rayyan, a web based screening application.[@ouzzani2016]
Title and abstract screening to remove obviously irrelevant records will be performed by the primary author, with a random selection of excluded records being screened in duplicate to ensure consistency with the inclusion criteria. If this demonstrates a significant level of erroneous exclusion by the primary author a larger proportion will be dual-screened. 
Full-text screening will also be completed in full by the primary author. A second reviewer will screen a random sample of included and excluded records, in addition to any records identified by the first reviewer as being difficult to assess against the inclusion criteria. Reasons for exclusion at this stage will be recorded. Disagreements occurring during either stage of the screening process will be resolved through discussion with a senior colleague. A PRIMSA flow diagram will be produced to document how records moved through the review.[@zotero-766]

#### Inclusion
We will seek studies that examine the relationship between blood lipid levels (or any specific lipid fraction, including total cholesterol, HDL, LDL, and triglycerides) and risk of incident dementia/MCI. Eligible study designs include randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Participants will be free (or assumed to be free) of dementia/MCI at baseline. Studies of any duration will be included to allow for exploration of the effect of length of follow-up on the effect estimate using meta-regression. No limits will be placed on the sample size of included studies.

Eligible studies will define dementia according to recognised criteria, for example the National Institute of Neurological Disorders and Stroke Association-Internationale pour la Recherche en l’Enseignement en Neurosciences (NINDS-AIREN), International Classification of Diseases (ICD), or Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria. For MCI, eligible studies are those that attempted state a definition for diagnoses of MCI (e.g. an adapted version of the Petersen criteria [@petersen1999]) and create ordinal groups of patients (e.g. no dementia or dementia/MCI/dementia) based on this definition.

No limitations will be imposed on publication status, publication date, venue or language, although we will require sufficiently detailed reports of the studies to be able to examine their methods. Preprints and unpublished reports will be eligible for inclusion if relevant. Multiple publications resulting from the analysis of the same data will be included and grouped.

#### Exclusion
Case-control studies, cross-sectional studies, qualitative studies, case reports/series and narrative reviews will be excluded. Studies which present no evidence of attempting to exclude prevalent cases from their analyses will also be excluded. Studies that measure change in continuous cognitive measures (e.g. MoCA score) without attempt to map these scores to ordinal groups (e.g. no dementia/MCI/dementia) will be excluded. Conference abstracts with no corresponding full-text publication will be examined, and we will contact authors to obtain information on the study’s status. Studies that are reported in insufficient detail (e.g. only in conference abstracts, new, letters, editorials and opinion) will be excluded. Previous systematic reviews will not be eligible, but their reference lists will be screened to identify any potentially relevant articles. Studies with outcomes not directly related to the clinical syndrome of dementia (e.g., neuroimaging), studies implementing a “multidomain intervention” where the lipid-regulating agent is included in each arms (e.g. for example, a study examining exercise + statins vs statins alone, but a study examining exercise + statins vs exercise alone would be included), and studies where there was no screening for dementia at baseline except if the sample was initially assessed in mid-life (i.e. below the age of 50) will be excluded.

### Data extraction

### Risk of bias assessment

Risk of bias assessment was performed using the domain-based risk-of-bias assessment tool appropriate to the study design. Randomised controlled trials were assessed using the RoB2 tool,[@sterne2019], non-randomised studies of interventions were assessed using the ROBINS-I tool,[@sterne2016] and non-randomised studies of exposures were assessed using the ROBINS-E tool.[TBC]

At present, no risk of bias assessment tool for Mendelian randomisation studies is available. Bias in thesee studies was assessed with the help of an expert panel [TBC]

### Patient and public involvement

## Results

```{r prisma-flow, echo = FALSE, warning=FALSE}
```

## Discussion

## Conclusion

### 2020-02-07 docs/5abd0cd3e96bf9459398f146495c15129ee56682/_main.tex
Get PDF download working
Major updates

This is a test of the bibliography (\protect\hyperlink{ref-Darwin1859}{1})

\begin{savequote}
There is grandeur in this view of life, with its several powers, having
been originally breathed into a few forms or into one; and that, whilst
this planet has gone cycling on according to the fixed law of gravity,
from so simple a beginning endless forms most beautiful and most
wonderful have been, and are being, evolved.
\qauthor{--- Charles Darwin\textsuperscript{\protect\hyperlink{ref-Darwin1859}{1}}}\end{savequote}

@@

\hypertarget{test}{%
\section{Test}\label{test}}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Darwin1859}{}%
1. Darwin, C. \emph{On the Origin of Species by Means of Natural Selection or the Preservation of Favoured Races in the Struggle for Life}. (John Murray, 1859).

>>>>

This is a test of the bibliography\textsuperscript{{\textbf{???}}}

The following tools were used to perform risk of bias assessments: ROB-2, ROBINS-I and QUADAS-2

\begin{savequote}
Hold
\qauthor{--- Charles Darwin\textsuperscript{{\textbf{???}}}}\end{savequote}

@@

\hypertarget{aims}{%
\section{Aims}\label{aims}}

The aim of this chapter is to systematically review all available literature on the association between blood levels of total cholesterol and it's constituent parts (HDL-c,LDL-c and triglycerides) on the subsequent risk of dementia.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{search-strategy}{%
\subsection{Search strategy}\label{search-strategy}}

We will systematically search electronic bibliographic databases to identify potentially relevant records. The search strategy used in each database will be developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE) terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy will include terms related to lipids, lipid modifying treatments, and dementia and its sub-types, and will be designed for MEDLINE before being adapted for use in the other bibliography databases listed. An outline of the general strategy is presented in the Table 3.2 below and the full draft search strategies for each database are attached to this protocol. To ensure that the study design filters are not excluding potentially relevant records, a random sample of 500 records identified by the main search but excluded by the filters (defined as Line 7 NOT Line 13 in Table 3.2) will be screened. If any potentially relevant studies are identified, their titles and abstracts will be searched for key terms that can be incorporated into the filters to improve search sensitivity.

The following databases will be searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection. We will also search clinical trial registries, for example ClinicalTrials.gov, to identify relevant randomized controlled trials.

The abstracts list of relevant conferences (e.g.~the proceedings of the Alzheimer's Association International Conference, published in the journal Alzheimer's \& Dementia) will be searched. Grey literature will also be searched via ProQuest, OpenGrey and Web of Science Conference Proceedings Citation Index, while theses will be accessed using the Open Access Theses and Dissertations portal. We will also search BioArxiv, a preprint repository, to identify potentially relevant studies. Finally, the reference lists of included studies will be searched by hand while studies citing included studies will be examined using Google Scholar (forward and reverse citation searching).

\hypertarget{study-selection}{%
\subsection{Study selection}\label{study-selection}}

Records will be imported into Endnote and deduplicated using the method outlined in Bramer et al.~(2016).\textsuperscript{\protect\hyperlink{ref-bramer2016}{1}} Screening (both title/abstract and full-text) will be performed using a combination of Endnote and Rayyan, a web based screening application.\textsuperscript{\protect\hyperlink{ref-ouzzani2016}{2}}
Title and abstract screening to remove obviously irrelevant records will be performed by the primary author, with a random selection of excluded records being screened in duplicate to ensure consistency with the inclusion criteria. If this demonstrates a significant level of erroneous exclusion by the primary author a larger proportion will be dual-screened.
Full-text screening will also be completed in full by the primary author. A second reviewer will screen a random sample of included and excluded records, in addition to any records identified by the first reviewer as being difficult to assess against the inclusion criteria. Reasons for exclusion at this stage will be recorded. Disagreements occurring during either stage of the screening process will be resolved through discussion with a senior colleague. A PRIMSA flow diagram will be produced to document how records moved through the review.\textsuperscript{\protect\hyperlink{ref-zotero-766}{3}}

\hypertarget{inclusion}{%
\subsubsection{Inclusion}\label{inclusion}}

We will seek studies that examine the relationship between blood lipid levels (or any specific lipid fraction, including total cholesterol, HDL, LDL, and triglycerides) and risk of incident dementia/MCI. Eligible study designs include randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Participants will be free (or assumed to be free) of dementia/MCI at baseline. Studies of any duration will be included to allow for exploration of the effect of length of follow-up on the effect estimate using meta-regression. No limits will be placed on the sample size of included studies.

Eligible studies will define dementia according to recognised criteria, for example the National Institute of Neurological Disorders and Stroke Association-Internationale pour la Recherche en l'Enseignement en Neurosciences (NINDS-AIREN), International Classification of Diseases (ICD), or Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria. For MCI, eligible studies are those that attempted state a definition for diagnoses of MCI (e.g.~an adapted version of the Petersen criteria\textsuperscript{\protect\hyperlink{ref-petersen1999}{4}}) and create ordinal groups of patients (e.g.~no dementia or dementia/MCI/dementia) based on this definition.

No limitations will be imposed on publication status, publication date, venue or language, although we will require sufficiently detailed reports of the studies to be able to examine their methods. Preprints and unpublished reports will be eligible for inclusion if relevant. Multiple publications resulting from the analysis of the same data will be included and grouped.

\hypertarget{exclusion}{%
\subsubsection{Exclusion}\label{exclusion}}

Case-control studies, cross-sectional studies, qualitative studies, case reports/series and narrative reviews will be excluded. Studies which present no evidence of attempting to exclude prevalent cases from their analyses will also be excluded. Studies that measure change in continuous cognitive measures (e.g.~MoCA score) without attempt to map these scores to ordinal groups (e.g.~no dementia/MCI/dementia) will be excluded. Conference abstracts with no corresponding full-text publication will be examined, and we will contact authors to obtain information on the study's status. Studies that are reported in insufficient detail (e.g.~only in conference abstracts, new, letters, editorials and opinion) will be excluded. Previous systematic reviews will not be eligible, but their reference lists will be screened to identify any potentially relevant articles. Studies with outcomes not directly related to the clinical syndrome of dementia (e.g., neuroimaging), studies implementing a ``multidomain intervention'' where the lipid-regulating agent is included in each arms (e.g.~for example, a study examining exercise + statins vs statins alone, but a study examining exercise + statins vs exercise alone would be included), and studies where there was no screening for dementia at baseline except if the sample was initially assessed in mid-life (i.e.~below the age of 50) will be excluded.

\hypertarget{data-extraction}{%
\subsection{Data extraction}\label{data-extraction}}

\hypertarget{risk-of-bias-assessment}{%
\subsection{Risk of bias assessment}\label{risk-of-bias-assessment}}

Risk of bias assessment was performed using the domain-based risk-of-bias assessment tool appropriate to the study design. Randomised controlled trials were assessed using the RoB2 tool,\textsuperscript{\protect\hyperlink{ref-sterne2019}{5}}, non-randomised studies of interventions were assessed using the ROBINS-I tool,\textsuperscript{\protect\hyperlink{ref-sterne2016}{6}} and non-randomised studies of exposures were assessed using the ROBINS-E tool.{[}TBC{]}

At present, no risk of bias assessment tool for Mendelian randomisation studies is available. Bias in thesee studies was assessed with the help of an expert panel {[}TBC{]}

\hypertarget{patient-and-public-involvement}{%
\subsection{Patient and public involvement}\label{patient-and-public-involvement}}

\hypertarget{results}{%
\section{Results}\label{results}}

\includegraphics{_main_files/figure-latex/prisma-flow-1.pdf}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{annotated-bibliography}{%
\chapter{Annotated Bibliography}\label{annotated-bibliography}}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-bramer2016}{}%
1. Bramer, W. M., Giustini, D., de Jonge, G. B., Holland, L. \& Bekhuis, T. De-duplication of database search results for systematic reviews in EndNote. \emph{Journal of the Medical Library Association : JMLA} \textbf{104}, 240--243 (2016).

\leavevmode\hypertarget{ref-ouzzani2016}{}%
2. Ouzzani, M., Hammady, H., Fedorowicz, Z. \& Elmagarmid, A. Rayyan-a web and mobile app for systematic reviews. \emph{Systematic Reviews} \textbf{5}, 210 (2016).

\leavevmode\hypertarget{ref-zotero-766}{}%
3. The PRISMA statement for reporting systematic reviews and meta-analyses of studies that evaluate healthcare interventions: Explanation and elaboration \textbar{} The BMJ.

\leavevmode\hypertarget{ref-petersen1999}{}%
4. Petersen, R. C. \emph{et al.} Mild cognitive impairment: Clinical characterization and outcome. \emph{Archives of Neurology} \textbf{56}, 303--308 (1999).

\leavevmode\hypertarget{ref-sterne2019}{}%
5. Sterne, J. A. C. \emph{et al.} RoB 2: A revised tool for assessing risk of bias in randomised trials. \emph{BMJ} \textbf{366}, (2019).

\leavevmode\hypertarget{ref-sterne2016}{}%
6. Sterne, J. A. \emph{et al.} ROBINS-I: A tool for assessing risk of bias in non-randomised studies of interventions. \emph{BMJ} \textbf{355}, (2016).

### 2020-02-07 index.Rmd
Major updates

#####################
## output options  ##
#####################
output:
  bookdown::pdf_book:
    template: templates/template.tex
    keep_tex: true
    citation_package: none  
    pandoc_args: ["--lua-filter=scripts_and_filters/correction_filter.lua", "--csl","bibliography/nature.csl"] #remove filter to stop applying blue background to inline corrections
  bookdown::gitbook:
    css: templates/style.css
    download: ["pdf"]
    config:
      toc:
        before: |
          <li><a href="./">Thesis - Draft</a></li>
        after: |
          <li><a href="https://github.com/mcguinlu/thesis" target="blank">Associated repository</a></li>
      sharing:
        facebook: false
        twitter: yes
        all: false
  bookdown::word_document2:
    toc: true   
link-citations: true
documentclass: book

>>>>

#####################
## output options  ##
#####################
output:
  bookdown::pdf_book:
    template: templates/template.tex
    keep_tex: true
    citation_package: none  
    pandoc_args: ["--lua-filter=scripts_and_filters/correction_filter.lua", "--csl","bibliography/nature.csl"] #remove filter to stop applying blue background to inline corrections
  bookdown::gitbook:
    css: templates/style.css
    download: ["pdf"]
    split_bib: false
    config:
      toc:
        before: |
          <li><a href="./">Thesis - Draft</a></li>
        after: |
          <li><a href="https://github.com/mcguinlu/thesis" target="blank">Associated repository</a></li>

      sharing:
        facebook: false
        twitter: yes
        all: false
  bookdown::word_document2:
    toc: true   
link-citations: true
documentclass: book
